---
title: "DNA Prevention Analysis"
author: "Bernard Lee"
date: '2020-04-03'
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: inline
---

#### Load library
```{r}
library(edgeR)
library(tidyverse)
library(factoextra)
library(M3C)
library(COMBAT)
library(sva)
#library(readxl)
#library(sva)
#library(bladderbatch)

#library(ape)
#library(ggpubr)
#library(gridExtra)
#library(grid)
```

#### Read in DNA prevention gene expression and metadata
```{r}
# Read in DNA prevention gene expression data
d0 <- read.delim("~/OneDrive - Cancer Research Malaysia/Bryan/2020/Bryan/Data/WC/51_DNA_Prevention_Gene_Counts.txt", row.names = 1, check.names = FALSE)

# Metadata
metadata <- as.data.frame(colnames(d0))
names(metadata) <- "Sample"
d1 <- read.delim("~/OneDrive - Cancer Research Malaysia/Bryan/2020/Bryan/Data/WC/51_samples_metadata.txt")
metadata <- left_join(metadata, d1)
metadata$Sample <- as.character(metadata$Sample)
metadata$Batch <- as.numeric(as.character(metadata$Batch))

# Multidimensional scaling (MDS) plot
plotMDS(d0, col = as.numeric(metadata$Disease))

# Remove data with zero counts
cutoff <- 1
drop <- which(apply(cpm(d0), 1, max) < cutoff)
d <- d0[-drop,] 
dim(d) # number of genes left

# Filter low expressed genes
d0 <- DGEList(d0)
keep.exprs <- filterByExpr(d0)
d0 <- d0[keep.exprs, keep.lib.sizes = FALSE]
summary(keep.exprs)

# Plot DNA prevention sample library size
barplot(d0$samples$lib.size, names.arg = colnames(d0), las=2, col = metadata$Colour, cex.names = 0.82)

# Adjusting for batch effects with Combat
d_mat <- as.matrix(d0)
batch = metadata$Batch
modcombat = model.matrix(~1, data = metadata)
combat_d0 = ComBat(dat = d_mat, batch = batch, mod = modcombat, par.prior=TRUE, prior.plots=FALSE)

# UMAP using COMBAT correct data
par(pty = "s")
umap(combat_d0, labels = as.factor(metadata$Disease), legendtitle = "Disease Stage")
#umap(combat_d0, labels = as.factor(metadata$Disease), legendtitle = "Disease Stage", text = metadata$Number)

# tsne using COMBAT corrected data
tsne(combat_d0, labels = as.factor(metadata$Disease), legendtitle = "Disease Stage")
#tsne(combat_d0, labels = as.factor(metadata$Disease), legendtitle = "Disease Stage", text = metadata$Number)
```

#### Normalisation
```{r}
d0 <- calcNormFactors(d0, method = "TMM")
tmm <- cpm(d0, log = TRUE)
write.table(tmm, "/Users/mac/OneDrive - Cancer Research Malaysia/Kuepeng/2020/New_targets/OPMD/51_DNA_Prevention_Gene_Counts_TMM_Normalized.txt", quote = FALSE, sep = "\t")
```

#### Design matrix using metadata file
```{r}
# Design matrix with the metadata file (For batch correction)
f <- factor(metadata$Disease)
batch <- factor(metadata$Batch)
design <- model.matrix(~0 + f + batch)
colnames(design) <- c("FEP","Mod_Severe","OSCC","Batch")
design
```

#### Calculate dispersion
```{r}
y <- estimateDisp(d0, design, robust = TRUE)
y$common.dispersion
#plotBCV(y)
```

#### Quasi-likehood dispersion
```{r}
fit <- glmQLFit(y, design, robust = TRUE)
head(fit$coefficients)
#plotQLDisp(fit)
```

#### Differential gene expression
#### Change the differential analysis phenotype comparison accordingly: 1. Mod_Severe -  FEP; 2. OSCC - FEP; 3. OSCC - Mod_Severe
```{r}
con <- makeContrasts(OSCC - Mod_Severe, levels = design)
qlf <- glmQLFTest(fit, contrast = con)
topTags(qlf)
```

#### TREAT
```{r}
tr <- glmTreat(fit, contrast = con, lfc = log2(1.2))
topTags(tr)
is.de <- decideTestsDGE(tr, adjust.method = "BH")
summary(is.de)
```

#### Write out the output
```{r}
tab <- topTags(tr, n = Inf, adjust.method = "BH")
write.table(tab$table, file="~/Desktop/Bryan/Output/DEG/_DNA_prevention_OSCC_vs_ModSevere.txt", sep = '\t', quote = FALSE, row.names = TRUE)
```

#### DEG analysis
```{r}
# Read in the files
GSE <- read.delim("~/Desktop/Bryan/Output/DEG/_DNA_prevention_OSCC_vs_ModSevere.txt", header = TRUE)
head(GSE)
GSE$lgtranspvalue <- -log10(GSE$PValue)
head(GSE)

par(pch = 16)
par(pty = "s")

# Plotting the figures
plot(GSE$logFC, GSE$lgtranspvalue, main='OSCC (n = 6) vs Mod-Severe (n = 31)', xlab=expression('lg'[2]*'FC'), ylab=expression('-lg'[10]*'pvalue'),cex.lab=1.2)
with(subset(GSE, logFC < 0 & lgtranspvalue >= (2/-logFC)), points(logFC,lgtranspvalue, col = "red"))
with(subset(GSE, logFC > 0 & lgtranspvalue >= (2/logFC)), points(logFC,lgtranspvalue, col = "blue"))

# Add legend
#legend("bottomright",legend=c(expression(paste('DOWN: lg'[2]*'FC<0 & -lg'[10]*'pvalue>=(2/-lg'[2]*'FC)')), expression(paste('UP: lg'[2]*'FC>0 & -lg'[10]*'pvalue>=(2/lg'[2]*'FC)'))),pch = 16, col=c("red", "blue"))

# Draw lines -------------------------------------------------------------------
xpos <- seq(0, 3, 0.01)
xneg <- seq(-5, 0, 0.01)
points(xpos, 2/xpos, type="l")
points(xneg, -2/xneg, type="l")

# Save down-regulated genes ----------------------------------------------------
GSE_DOWN <- subset(GSE, logFC < 0 & -log10(PValue)>=(2/-logFC) , select=c(logFC, PValue))
GSE_DOWN <- GSE_DOWN[order(GSE_DOWN$logFC), ]
nrow(GSE_DOWN)
GSE_DOWN$GeneName <- rownames(GSE_DOWN)
GSE_DOWN <- GSE_DOWN[, c(3, 1:2)]
head(GSE_DOWN)
write.table(GSE_DOWN, "~/Desktop/Bryan/Output/DEG/_DNA_prevention_OSCC_vs_ModSevere_DOWN.txt", quote = FALSE, sep = '\t', row.names = FALSE)

# Save up-regulated genes ------------------------------------------------------
GSE_UP <- subset(GSE, logFC > 0 & -log10(PValue)>=(2/logFC) , select=c(logFC, PValue))
GSE_UP <- GSE_UP[order(GSE_UP$logFC, decreasing = TRUE), ]
nrow(GSE_UP)
GSE_UP$GeneName <- rownames(GSE_UP)
GSE_UP <- GSE_UP[, c(3, 1:2)]
head(GSE_UP)
write.table(GSE_UP, "~/Desktop/Bryan/Output/DEG/_DNA_prevention_OSCC_vs_ModSevere_UP.txt", quote = FALSE, sep = '\t', row.names = FALSE)
```

#### GNGT1
```{r}
dat <- read.delim("~/Desktop/Bryan/Data/WC/GNGT1.txt", header = TRUE, check.names = FALSE)
head(dat)
ggplot(dat, aes(x = Disease_Stage, y = GNGT1)) +
  geom_boxplot() +
  geom_jitter(shape=16, position=position_jitter(0.2)) +
  geom_signif(comparisons = list(c("Mod-Severe", "FEP")), map_signif_level=TRUE) +
  labs(y = "GNGT1 expression", x = " Disease Stage") +
  theme_bw() +
  theme(aspect.ratio = 1)
ggsave("~/Desktop/Bryan/Figure/GNGT1.pdf")
```




#### Session information
```{r}
sessionInfo()
```

