---
title: "OPMD WGCNA analysis"
author: "Hui Shi Saw"
date: '2020-07-17'
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: inline
---

#### 1 Assembly and preprocessing of RNAseq data
```{r}
# Load library
library(tidyverse)
library(rafalib)
library(limma)
library(WGCNA)
cor <- WGCNA::cor
options(stringsAsFactors = FALSE)

# Load gene expression data
RNA_seq <- read.delim("C:/Users/saw hui shi/Desktop/CRMY/R Programming/OPMD/51_Samples_Counts.txt", header = TRUE, row.names = 1, check.names = FALSE)
dim(RNA_seq)
names(RNA_seq)

# Need to do some pre-processing to remove genes that are not expressed in more than 20% of the 51 samples
OPMD_rnaseq <- RNA_seq[apply(RNA_seq, 1, function(x) sum(x == 0)) < ncol(RNA_seq) * 0.8, ]

# Run voom
RNAseq_voom <- voom(OPMD_rnaseq)$E

# transpose matrix to correlate genes
WGCNA_matrix <- t(RNAseq_voom[order(apply(RNAseq_voom, 1, mad), decreasing = TRUE), ])

# Check for genes and samples with too many missing values
gsg <- goodSamplesGenes(WGCNA_matrix, verbose = 5)
gsg$allOK
```

#### Sample clustering
```{r}
sampleTree <- hclust(dist(WGCNA_matrix), method = "average")
par(cex = 0.6)
plot(sampleTree, main = "Sample clustering to detect outliers", sub = "", xlab = "", cex.lab = 1.5,
cex.axis = 1.5, cex.main = 2)
```

#### (cancelled) Load clinical trait data
```{r}
# Load clinical trait data
traitData = read.delim("/Users/mac/Desktop/OPMD/Data/51_samples_metadata_edited.txt")
dim(traitData)
names(traitData)

# Form a data frame analogous to expression data that will hold the clinical trait
clinicalSamples = rownames(WGCNA_matrix)
traitRows = match(clinicalSamples, traitData$Sample)
datTraits = traitData[traitRows, ]
rownames(datTraits) = traitData[traitRows, 1]
collectGarbage()
```

#### (cancelled) Visualize how the clinical traits (grading) relate to the sample dendrogram
```{r}
# Re-cluster samples
sampleTree2 <- hclust(dist(WGCNA_matrix), method = "average")

# Convert traits to a color representation: white means low, red means high, grey means missing entry
traitColors <- numbers2colors(as.fumeric(datTraits$Grading), signed = FALSE)

# Plot the sample dendrogram and the colors underneath.
plotDendroAndColors(sampleTree2, traitColors, groupLabels = "Grading", main = "Sample dendrogram and trait heatmap")

# Save data
save(WGCNA_matrix, datTraits, file = "/Users/mac/Desktop/OPMD/Library/51_OPMD_01-dataInput.RData")
```


#### 2 Automatic network construction and module detection
```{r}
# Choose a set of soft-thresholding powers
powers <- c(c(1:10), seq(from = 12, to = 20, by = 2))


# Call the network topology analysis function
sft <- pickSoftThreshold(WGCNA_matrix, powerVector = powers, verbose = 5)

# Power estimate
sft$powerEstimate
sft$fitIndices

# Plot the results:
par(mfrow = c(1,2))
cex1 <- 0.86

# Scale-free topology fit index as a function of the soft-thresholding power
plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2], xlab = "Soft Threshold (power)", ylab = "Scale Free Topology Model Fit,signed R^2", type = "n", main = paste("Scale independence"))
text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2], labels = powers, cex = cex1, col="red")
# this line corresponds to using an R^2 cut-off of h
abline(h = 0.86, col = "red")

# Mean connectivity as a function of the soft-thresholding power
plot(sft$fitIndices[,1], sft$fitIndices[,5], xlab = "Soft Threshold (power)", ylab = "Mean Connectivity", type="n", main = paste("Mean connectivity"))
text(sft$fitIndices[,1], sft$fitIndices[,5], labels = powers, cex = cex1, col="red")
```
#### Constructing the gene network and identifying modules
```{r}
net <- blockwiseModules(WGCNA_matrix, power = 5, TOMType = "unsigned", minModuleSize = 30, reassignThreshold = 0, mergeCutHeight = 0.25, numericLabels = TRUE, pamRespectsDendro = FALSE, saveTOMs = TRUE, saveTOMFileBase = "OPMDTOM", verbose = 5, maxBlockSize = 20000)

# no. of modules identified
table(net$colors)
```

#### Display dendrogram
```{r}
# Convert labels to colors for plotting
mergedColors <- labels2colors(net$colors)

# Plot the dendrogram and the module colors underneath
plotDendroAndColors(net$dendrograms[[1]], mergedColors[net$blockGenes[[1]]], "Module colors", dendroLabels = FALSE, hang = 0.03, addGuide = TRUE, guideHang = 0.05)
```

#### Save the module assignment and module eigengene information
```{r}
moduleLabels <- net$colors
moduleColors <- labels2colors(net$colors)
MEs <- net$MEs
geneTree <- net$dendrograms[[1]]
save(MEs, moduleLabels, moduleColors, geneTree, file = "/Users/mac/Desktop/OPMD/Library/51_OPMD_NetworkConstruction_Auto.RData")
```

#### (cancelled) Quantifying module–trait associations
```{r}
# Define numbers of genes and samples
nGenes <- ncol(WGCNA_matrix)
nSamples <- nrow(WGCNA_matrix)

# Recalculate MEs with color labels
MEs0 <- moduleEigengenes(WGCNA_matrix, moduleColors)$eigengenes
MEs <- orderMEs(MEs0)
moduleTraitCor <- cor(MEs, as.fumeric(datTraits$Grading), use = "p")
moduleTraitPvalue = corPvalueStudent(moduleTraitCor, nSamples)
```

#### (cancelled) Starts here
```{r}
load("51_OPMD-Auto_step.RData")
sizeGrWindow(10,6)
# Will display correlations and their p-values
textMatrix = paste(signif(moduleTraitCor, 2), "\n(",
signif(moduleTraitPvalue, 1), ")", sep = "");
dim(textMatrix) = dim(moduleTraitCor)
par(mar = c(6, 8.5, 3, 3));

# Display the correlation values within a heatmap plot
labeledHeatmap(Matrix = moduleTraitCor,
xLabels = names(datTraits),
yLabels = names(MEs),
ySymbols = names(MEs),
#### colorLabels = FALSE,
colors = greenWhiteRed(50),
textMatrix = textMatrix,
setStdMargins = FALSE,
cex.text = 0.5,
zlim = c(-1,1),
main = paste("Module-trait relationships"))
```

#### Get module genes
# Check genes in each module
```{r}
table(net$colors)

# Change the numerical and directory accordingly
# This is for module 1
module_Genes <- which(net$colors == 1)
module_Genes
write.table(module_Genes, "C:/Users/saw hui shi/Desktop/CRMY/R Programming/OPMD/moduleGenes_1.txt", sep = "\t", quote = FALSE)
```

#### Extract Gene Expression to construct heatmap
```{r}
library(tidyverse)

# Load module genes_30
dat <- read_delim("C:/Users/saw hui shi/Desktop/CRMY/R Programming/OPMD/Module_GeneExpression/moduleGenes_30_Name.txt", delim = "\t")

# Extract gene expression
geneExpression <- RNAseq_voom[which(rownames(RNAseq_voom) %in% dat$GeneName), ]

# Write the gene expression
write.table(geneExpression, "C:/Users/saw hui shi/Desktop/CRMY/R Programming/OPMD/Module_GeneExpression/module30_GeneExpression.txt", sep = "\t", quote = FALSE)
```


